---
title: "New panel functions"
subtitle: "tactile"
author: "Johan Larsson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{New panel functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 4,
  fig.height = 3.2
)
```

**tactile** introduces new panel functions to the latticeverse.

## `panel.ci()`: confidence intervals

`panel.ci()` adds confidence bands around a line using 
arguments `lower` and `upper`. This is usually of interested after,
for instance, having fitted a model and then made predictions using that 
model.

As an example, we will try to predict petal width from petal length
and species, using the `iris` dataset.

```{r model}
mod <- lm(Petal.Width ~ Petal.Length * Species, data = iris)
newdat <- expand.grid(
  Petal.Length = seq(1, 7, by = 0.1),
  Species = c("setosa", "versicolor", "virginica")
)
pred <- predict(mod, newdat, interval = "confidence")
dd <- cbind(newdat, pred)
```

Having predicted values across our grid, we now plot the predictions,
including 95% confidence levels using the following lines.

```{r panel-ci, fig.cap="Example of using panel.ci."}
library(tactile)
xyplot(fit ~ Petal.Length,
  groups = Species, data = dd,
  prepanel = prepanel.ci, auto.key = list(lines = TRUE, points = FALSE),
  ylab = "Petal Width",
  xlab = "Petal Length",
  lower = dd$lwr,
  upper = dd$upr,
  type = "l",
  panel = function(...) {
    panel.ci(..., alpha = 0.15, grid = TRUE)
    panel.xyplot(...)
  }
)
```

Also note the use of the `prepanel.ci()` function that we provide the
`prepanel` argument with so that the axis limits are set properly.

## `panel.qqmathci()`: confidence intervals for `lattice::qqmath()`

`panel.qqmathci()` is designed to be used together with `lattice::qqmath()`
and `lattice::panel.qqmathline()` to provide confidence
intervals for the theoretical quantities. A rather contrived example follows.

```{r qqmath, fig.cap="Example of using panel.qqmathci."}
qqmath(~ height | voice.part,
  aspect = "xy", data = singer,
  prepanel = prepanel.qqmathline,
  distribution = qnorm,
  ci = 0.9,
  panel = function(x, ...) {
    panel.qqmathci(x, ...)
    panel.qqmathline(x, ...)
    panel.qqmath(x, ...)
  }
)
```

